Formalizing the Kolmogorov extension Theorem in Lean

### 2023-09-12 Manuscript/proof_outline.tex
semiring -> semi-ring

\begin{lemma}\label{l1}
  Let $\mathcal H$ be a semiring, $\mathcal I \subseteq_f \mathcal H$,
  $A \in \mathcal H$. Then, there is $\mathcal K \subseteq_f \mathcal
  H$ such that $\mathcal K$ contains pairwise disjoint sets and $A
  \setminus \bigcup_{I \in \mathcal I} I = \biguplus_{K\in \mathcal K}
  K$.
\end{lemma}

\begin{proof}
  We proceed by induction on $\mathcal I$. If $\mathcal I$ is a
  singleton, the assertion is true by the definition of a semiring. If
  it holds for some $\mathcal I$ (i.e.\ there is $\mathcal K
  \subseteq_f \mathcal H$ with $A \setminus \bigcup_{I \in \mathcal I}
  I = \biguplus_{K\in \mathcal K} K$), let us consider $\mathcal I' =
  \{J\} \cup \mathcal I$ for some $J \notin \mathcal I$. For each $K
  \in \mathcal K$, Write $K \setminus J = \biguplus_{J_K \in \mathcal
    J_K} J_K$ for some $\mathcal J_K \subseteq_f \mathcal H$ (which
  exists by the definition of a semi-ring). Then, write
  \begin{align*}
    A \setminus \bigcup_{I' \in \mathcal I'} I & = \Big(A \setminus
    \bigcup_{I \in \mathcal I} I\Big) \setminus J = \biguplus_{K\in
      \mathcal K} K \setminus J = \biguplus_{K\in \mathcal K}
    \biguplus_{J_K \in \mathcal J_K} J_K.
  \end{align*}
  This concludes the proof, since the latter disjoint union is over a
  finite set.
\end{proof}

@@

For the concrete application we have in mind, we introduce a definition which uses a specific semiring, the measurable cylinders, based on a projective family of measures.
That is the goal of the two definitions \leanline{kolmogorovFun} and \leanline{kolContent} below:

@@

\noindent
Although some material on outer measures was covered in {\tt mathlib}
already, the classical extension theorem (extending a set function $m$
on a semiring $\mathcal H$)) was not provided yet. In particular, this
result states that the outer measure coincides with $m$ on $\mathcal
H$. All statements on equality of $\mu$ and $m$ present in {\tt
  mathlib} at the time of writing have too many requirements: they all require $m$ to be a $\sigma$-additive function defined on a $\sigma$-algebra.

%The next result talks about an \leanline{induced_outer_measure} in
%terms of {\tt mathlib}.
The next result extends a set function on a semi-ring\todo{check that we are consistend with semiring vs semi-ring} to an outer measure.

@@

\begin{theorem}[Carathéodory extension]\label{T:masseind}
  Let $\mathcal H$ be a semiring and $m: \mathcal H\to\mathbb R_+$
  $\sigma$-finite and $\sigma$-additive. Furthermore, let $\mu$ be the
  induced outer measure from Proposition~\ref{P:auss} and $\mathcal F$
  the $\sigma$-algebra from Theorem~\ref{T:cara}. Then,
  $\sigma(\mathcal H)\subseteq\mathcal F$ and $\mu$ coincides with $m$
  on $\mathcal H$.
\end{theorem}

@@

\begin{proof}
  1.$\Leftrightarrow$2.\ was already shown in
  Lemma~\ref{l:halbextRing}, since $\mathcal R$ is a semiring.

>>>>

\begin{lemma}\label{l1}
  Let $\mathcal H$ be a semi-ring, $\mathcal I \subseteq_f \mathcal H$,
  $A \in \mathcal H$. Then, there is $\mathcal K \subseteq_f \mathcal
  H$ such that $\mathcal K$ contains pairwise disjoint sets and $A
  \setminus \bigcup_{I \in \mathcal I} I = \biguplus_{K\in \mathcal K}
  K$.
\end{lemma}

\begin{proof}
  We proceed by induction on $\mathcal I$. If $\mathcal I$ is a
  singleton, the assertion is true by the definition of a semi-ring. If
  it holds for some $\mathcal I$ (i.e.\ there is $\mathcal K
  \subseteq_f \mathcal H$ with $A \setminus \bigcup_{I \in \mathcal I}
  I = \biguplus_{K\in \mathcal K} K$), let us consider $\mathcal I' =
  \{J\} \cup \mathcal I$ for some $J \notin \mathcal I$. For each $K
  \in \mathcal K$, Write $K \setminus J = \biguplus_{J_K \in \mathcal
    J_K} J_K$ for some $\mathcal J_K \subseteq_f \mathcal H$ (which
  exists by the definition of a semi-ring). Then, write
  \begin{align*}
    A \setminus \bigcup_{I' \in \mathcal I'} I & = \Big(A \setminus
    \bigcup_{I \in \mathcal I} I\Big) \setminus J = \biguplus_{K\in
      \mathcal K} K \setminus J = \biguplus_{K\in \mathcal K}
    \biguplus_{J_K \in \mathcal J_K} J_K.
  \end{align*}
  This concludes the proof, since the latter disjoint union is over a
  finite set.
\end{proof}

@@

For the concrete application we have in mind, we introduce a definition which uses a specific semi-ring, the measurable cylinders, based on a projective family of measures.
That is the goal of the two definitions \leanline{kolmogorovFun} and \leanline{kolContent} below:

@@

\noindent
Although some material on outer measures was covered in {\tt mathlib}
already, the classical extension theorem (extending a set function $m$
on a semi-ring $\mathcal H$)) was not provided yet. In particular, this
result states that the outer measure coincides with $m$ on $\mathcal
H$. All statements on equality of $\mu$ and $m$ present in {\tt
  mathlib} at the time of writing have too many requirements: they all require $m$ to be a $\sigma$-additive function defined on a $\sigma$-algebra.

%The next result talks about an \leanline{induced_outer_measure} in
%terms of {\tt mathlib}.
The next result extends a set function on a semi-ring to an outer measure.

@@

\begin{theorem}[Carathéodory extension]\label{T:masseind}
  Let $\mathcal H$ be a semi-ring and $m: \mathcal H\to\mathbb R_+$
  $\sigma$-finite and $\sigma$-additive. Furthermore, let $\mu$ be the
  induced outer measure from Proposition~\ref{P:auss} and $\mathcal F$
  the $\sigma$-algebra from Theorem~\ref{T:cara}. Then,
  $\sigma(\mathcal H)\subseteq\mathcal F$ and $\mu$ coincides with $m$
  on $\mathcal H$.
\end{theorem}

@@

\begin{proof}
  1.$\Leftrightarrow$2.\ was already shown in
  Lemma~\ref{l:halbextRing}, since $\mathcal R$ is a semi-ring.
